
```{r, include = FALSE, eval = FALSE}
ottrpal::set_knitr_image_path()
```


  
# Open Case Study Infrastructure

## Learning Objectives

In this chapter we will discuss the overall infrastructure of the DaSEH Project, which includes:

- [The DaSEH website](https://daseh.org/)
- Methods for learners to provide feedback
- [The DaSEH GitHub repository](https://github.com/fhdsl/DaSEH/)
- Our [R package called `dasehr`](https://github.com/fhdsl/dasehr)


## DaSEH Website



The [DaSEH website](https://daseh.org/) describes the mission of the DaSEH project, information for learners to apply to participate, a summary of how the course works, links to our learning content, and links to additional resources. 

Links to all of our learning modules can be found on the DaSEH website on the  on the [Materials Schedule page](https://daseh.org/materials_schedule.html). 

## Feedback

We are continually striving to make our content better. Please contact us if you have ideas for suggestions for the project.

Also please let us know if you notice typos or errors, or if you are interested in getting involved.

### Contact Info

The website contains a contact email that may be used to send a message to DaSEH to ask a question or provide suggestions.

### Survey

There is also a [survey](https://docs.google.com/forms/d/e/1FAIpQLSfpN4FN3KELqBNEgf2Atpi7Wy7Nqy2beSkFQINL7Y5sAMV5_w/viewform?usp=sf_link) available on our website that allows us to do research on DaSEH resource use. 

The survey should take no more than 10 minutes to complete. Your feedback helps us learn more about how to improve our resources. Part of this includes getting a better understanding of who is using our resources and how so that we can better design our materials. We would greatly appreciate you filling it out if you have the time!


## Materials Search Tool

The website also includes a Materials search tool to aid instructors in finding appropriate modules for their learning objectives. Accessing the search tool and how to use it is described in more detail below.

<!-- This diagram illustrates the workflow of accessing a case study from the OCS website through the case study search table. From the table, users can use the provided links to view the original static case studies, interactive case studies, and the GitHub repositories for each. Users may find all case study source files in the case study repository, as well as instructions on how to use the case study. -->



### Materials

We have the following modules on various topics related to data science for environmental public health. 



Each one has lecture slides built using ioslides as well as lab activities.

The rendered version of the slides, as well as the raw Rmd file are available on the website. In addition, the rendered and raw versions of the both the lab and the lab key are also available.

We also include cheatsheets so that learners can review the functions that they learned that day.

```{r, fig.align='center', echo = FALSE, fig.alt = "image of the materials and schedule table on the DaSEH website which shows the module name, the lecture links available in HTML, PDF, and the raw Rmd file, the lab activity links including the Rmd file and  html rendered version, as well as for the key version which includes the answers and finally additional resources like cheatsheets", out.width="100%"}
ottrpal::include_slide('https://docs.google.com/presentation/d/1vCiMPvvsdwQjiMWjf0YuSpTkG0DGXsy1614cRiFc7ns/edit?slide=id.g3a4b18ec263_0_0#slide=id.g3a4b18ec263_0_0')
```

1) An Introduction to R

This module includes the history of how R came to be and how it differs from other options like Python or Stata. It also includes how students should anticipate the experience of learning in the course and suggestions for how to learn. It also introduces jargon such as variable, row, column, object.

This module does not have a lab associated with it.

2) Basic R

This module includes how to assign objects, how to create vectors, and how to perform simple calculations.

The lab activity which is just an html webpage (as learners are not yet introduced to Rmd files) involves performing simple mathematical opperations, assigning vectors, and checking the class and length of vectors.

3) R Studio

This module includes a tour of R Studio, how to write code that can be saved or how to write interactive code, for instance for installing a package where the user may have to answer questions.

The lab activity involves knitting the Rmd file, running code chunks in the Rmd file, running all previous chunks, creating new chunks, and adding headers.

4) Reproducibility

This module introduces concepts related to repeatability, reproducibility, and replicability. It introduces best practices for improving transparency in our code.

The lab activity features tasks like cleaning the environment, using the `set.seed()` function to generate the same random numbers each time, and using the `sessionInfo()` function to list packages and versions used.

5) 


## Open Case Studies GitHub Organization

GitHub is a website and cloud service that enables developers to store, manage, and track changes to their code. OCS uses GitHub for both development and distribution purposes. Users have complete access to all case study material through our [OCS GitHub page](https://github.com/opencasestudies) where each case study is hosted in an individual repository. The repository contains all the materials needed for the case study. This includes the case study text to be distributed to students, the data used in the case study (discussed below), additional documents and references, and brief guidelines on case study use.

Data included in the GitHub repository is available in multiple formats to enable modular use of the case studies. This diagram explains the case study data folder structure and how data is categorized into different sub-folders:



Data included in the GitHub repository is available in multiple formats to facilitate modularization of the case studies as described below. To use the case study data, you can download the GitHub repository directly or use the `OCSdata` R package described below. 

## OCSdata 

To simplify the process of accessing the data required for each case study, we have created the `OCSdata` R package. Briefly, the `OCSdata` package creates a new folder called "OCSdata" where it downloads the data needed for a specific case study. Users can download the data in its original raw format or in various processed formats that correspond to different stages of data wrangling and cleaning. This allows users to perform the data exploration and wrangling or the data visualization and analysis sections of the case study without having to process the data from the raw files. For some of the case studies, the `OCSdata` package also downloads extra source data that is not used in the case study. 

The following are the main functions to import data in various formats using the `OCSdata` package. Each function is described in more detail in the `OCSdata` [package documentation](https://cran.r-project.org/web/packages/OCSdata/index.html).

| Data Folder | Case Study Section | `OCSdata` Function |
| ----- | -------- | ------ |
| raw | Data Import | `raw_data` |
| imported | Data Exploration, Data Wrangling | `imported_data` |
| wrangled | Data Visualization, Data Analysis | `wrangled_csv`, `wrangled_rda` |
| simpler_import | Data Import | `simpler_import_data` |
| extra | Suggested Homework (?) | `extra_data` |

The package source files and documentation are also available on [GitHub](https://github.com/opencasestudies/OCSdata).

### Getting Started with `OCSdata`

The `OCSdata` package is available on the package repository [CRAN](https://cran.r-project.org/web/packages/OCSdata/index.html). It requires R 3.5 or higher and can be installed in R as follows: 



### Downloading raw data

The `raw_data` function will download the raw data files that can be imported into R. 

The first argument is the name of the case study. A list of case study names can be found in the package documentation [online](https://cran.r-project.org/web/packages/OCSdata/vignettes/instructions.html#casestudy) or by typing `?raw_data` in R. 

The `outpath` argument is a string specifying the folder where the data should be downloaded. To download the data to a folder named "OCS_data" in the current working directory, you can supply `getwd()` to the `output` argument. If nothing is provided for the argument, you will be prompted to enter 1, 2, or 3 to download the data into the current director, to specify the download path, or to cancel, respectively. 

In the following example, we download the raw data for the "Opioids in the United States" case study to the current directory. 



### Downloading data in other formats

The `OCSdata` package can be used to download the data in various processed formats that may be helpful in skipping certain case study sections and focusing on data wrangling and/or analysis and visualization. All of the functions take the same arguments described above.

#### Simpler import

The `simpler_import_data()` function will download raw data files that have been converted to file formats that are easier to import into R, typically .csv. Some case studies offer this option when the original raw files require a more complicated import step. 



#### Importing data as R objects 

The `imported_data()` function will download raw data files in .rda format. This means the data have already been imported into R objects. This can be used to skip the data import section and start directly with data wrangling. The R objects files can be imported into R by either double clicking on the files in RStudio or using the `load()` function as follows. 




#### Importing wrangled data

The following functions will download the data files that have already been wrangled and are ready to be analyzed. These come in both .csv and .rda formats.

Download as csv files:



Downloading as R objects:



### Downloading extra data

Some case studies have extra data are not used in the case study but can be used to explore the case study subject from different perspectives. These data  but can This data can be downloaded using the `extra_data()` function. 



### Downloading all case study data

The `zip_ocs()` function will download the all of the repository files in a .zip folder and unzip them into a specified directory. This includes the case study data in all the formats detailed above (raw, simpler_import, imported, wrangled, and extra). It also includes the case study .Rmd file, which can be modified by instructors as needed. We recommend using this method over cloning or forking (terms that you may be familiar with if you are familiar with Git and GitHub), as this will not result in the user getting all of our git history. 

If you choose to fork the repository you will automatically generate a repository on GitHub and your repository will have connections to the original case study. This can be helpful for pulling any changes to the original case study. It can also be helpful if you want to send edits to the original case study in the form of what is called a pull request.  

If you clone the case study repository, you can set it up on GitHub as well with a few more steps and you will not preserve any connection to the original case study repository.

Again, don't worry if all these terms are new to you. You can just use the `zip_ocs()` function instead. Otherwise take a look at @happygitwithr to learn more.



### Fork or clone the case study repository

If instead users are familiar with Git and GitHub and want to fork or clone the case study repository, this can also easily be done using the `OCSdata` package. The `clone_ocs()` function of the `OCSdata` package can be used to do either. If the `fork_repo` function is set to `TRUE` it will fork the repo, otherwise, by default, it will clone the repository. These functions will result in the same outcome as using GitHub to clone or fork the repo.

Again you can also specify the `outpath` location as in the previous description about the `zip_ocs()` function.



However, using these functions will involve the users getting all of our git history so we suggest that users use the `zip_ocs()` function (described in the above section) of `OCSdata` instead. 


If you'd like to learn more about the `OCSdata` package or the OCS GitHub organization page, you can read this [thesis](https://jscholarship.library.jhu.edu/handle/1774.2/66820) by former graduate student [Michael Breshock](https://mbreshock.github.io/) @breshock_expanding_2021.

## Session info

```{r, echo = FALSE}
sessionInfo()
```
